---
title: "Technical Appendix"
author: '1632905'
date: "4 03 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

Loading required libraries.
```{r libraries, message = FALSE, warning=FALSE}
library(rio)
library(dplyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(gridExtra)
library(xts)
library(tidyr)
library(ggcorrplot)

```

Data was taken from different releases which need to be pre-processed individually because of slight differences in the structure. 

Pre-processign fo the older data.
```{r older data, message = FALSE}
#importing datasets

monthYears2 <- paste0(month.abb[c(11:12, 1:10)], c(rep(17, 2), rep(18,10)))
d2 <-list()
for  (i in 1:length(monthYears2)) {
  filename2 <- paste0("Appointments_GP_Oct18/CCG_CSV_", monthYears2[i], ".csv")
  d2[[i]] <-  import(filename2, setclass = "tibble")
}
covData17 <- tibble()
for (i in 1:length(monthYears2)) {
  covData17 <- rbind(covData17, filter(d2[[i]], CCG_NAME=="NHS Coventry and Rugby CCG"))
}
write.csv(covData17, "covData17.csv")


#selecting interesting features
covData17a <- select(covData17, CCG_CODE, APPOINTMENT_DATE, APPOINTMENT_STATUS, HCP_TYPE, APPOINTMENT_MODE,
                   TIME_BETWEEN_BOOK_AND_APPT, COUNT_OF_APPOINTMENTS)


# unifying the date format
for (i in 1:length(covData17a$APPOINTMENT_DATE)) {
  if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Jan"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/01/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Feb"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/02/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Mar"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/03/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Apr"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/04/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "May"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/05/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Jun"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/06/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Jul"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/07/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Aug"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/08/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Sep"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/09/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Oct"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/10/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Nov"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/11/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
  else if (substr(covData17a$APPOINTMENT_DATE[i], 4, 6) == "Dec"){
    covData17a$APPOINTMENT_DATE[i] <- paste0(substr(covData17a$APPOINTMENT_DATE[i], 0, 2), "/12/",20,  substr(covData17a$APPOINTMENT_DATE[i], 8, 11))
  }
}
covData17a$Weekday <- wday(dmy(covData17a$APPOINTMENT_DATE))
covData17a$Weekday <- recode(covData17a$Weekday, 
                                "1"="Sunday",
                                "2"="Monday",
                                "3"="Tuesday",
                                "4"="Wednesday",
                                "5"="Thursday",
                                "6"="Friday",
                                "7"="Saturday")

# summing the data over days
sumData17 <- aggregate(covData17a$COUNT_OF_APPOINTMENTS, by=list(Appointment_Date=covData17a$APPOINTMENT_DATE), FUN=sum)

# unifying th date format
sumData17$Weekday <- wday(dmy(sumData17$Appointment_Date))
sumData17$Weekday <- recode(sumData17$Weekday, 
                          "1"="Sunday",
                          "2"="Monday",
                          "3"="Tuesday",
                          "4"="Wednesday",
                          "5"="Thursday",
                          "6"="Friday",
                          "7"="Saturday")



# plotting the relationship between appointments and weekday
sumData17$Appointment_Date <- dmy(sumData17$Appointment_Date)
ggplot(sumData17, aes(x = Appointment_Date, y = x, color = Weekday)) + geom_jitter() + labs(y = "Visits per day", x = "Date")
Weekmean17 <- aggregate(sumData17$x, by=list(a = sumData17$Weekday), FUN = mean)



# importing the coverage dataset
coverage17 <- import("Appointments_GP_Oct18/APPOINTMENTS_GP_COVERAGE.csv")
coverage17 <- coverage17[which(coverage17$COMMISSIONER_ORGANISATION_CODE == "05A"),]

for (i in 1:length(coverage17$Appointment_Month)) {
  if (substr(coverage17$Appointment_Month[i], 4, 6) == "Jan"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/01/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Feb"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/02/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Mar"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/03/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Apr"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/04/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "May"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/05/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Jun"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/06/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Jul"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/07/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Aug"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/08/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Sep"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/09/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Oct"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/10/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Nov"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/11/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage17$Appointment_Month[i], 4, 6) == "Dec"){
    coverage17$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/12/",20,  substr(coverage17$Appointment_Month[i], 8, 11))
  }
}



# merging coverage with covData
covData17a$patients <- list(0)
covData17a$all_patients <- list(0)
covData17a$ratio_patients <- list(0)

for (i in 1:nrow(coverage17)) {
  for (j in 1:nrow(covData17a)) {
    if(substr(coverage17$Appointment_Month[i], 4, 11) == substr(covData17a$APPOINTMENT_DATE[j], 4, 11)) {
      covData17a$patients[j] <- coverage17$`Patients registered at included practices`[i]
      covData17a$all_patients[j] <- coverage17$`Patients registered at open practices`[i]
    }
      
  }
}
covData17a$ratio_patients <- as.numeric(covData17a$patients)/as.numeric(covData17a$all_patients)



# preparing covdatafor a merger
names(covData17a)[2] <- "Appointment_Date"
names(covData17a)[3] <- "APPT_STATUS"
names(covData17a)[5] <- "APPT_MODE"


#write covData17 as a csv
covData17a$patients <- unlist(covData17a$patients)
covData17a$all_patients <- unlist(covData17a$all_patients)

write.csv(covData17a, "old_data.csv", row.names = FALSE)




```

Pre-processing of the newer data.
```{r newer data, message = FALSE}
##importing the dataset from csv file

monthYears <- paste0(month.abb[c(3:12, 1:12)], "_", c(rep(18, 10), rep(19, 12)))
d <-list()
for  (i in 1:length(monthYears)) {
  filename <- paste0("Appointments_GP_Daily_Aug19/CCG_CSV_", monthYears[i], ".csv")
  d[[i]] <-  import(filename, setclass = "tibble")
}

##filetring dataset to get only data from COventry and Rugby
covData <- tibble()
for (i in 1:length(monthYears)) {
  covData <- rbind(covData, filter(d[[i]], CCG_NAME=="NHS Coventry and Rugby CCG"))
}
write.csv(covData, "covData.csv")


#selecting variables

covData1 <- select(covData, CCG_CODE, Appointment_Date, APPT_STATUS, HCP_TYPE, APPT_MODE,
                   TIME_BETWEEN_BOOK_AND_APPT, COUNT_OF_APPOINTMENTS)

# unifying the date format

for (i in 1:length(covData1$Appointment_Date)) {
  if (substr(covData1$Appointment_Date[i], 3, 5) == "JAN"){
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/01/", substr(covData1$Appointment_Date[i], 6, 9))
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "FEB") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/02/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "MAR") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/03/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "APR") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/04/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "MAY") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/05/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "JUN") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/06/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "JUL") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/07/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "AUG") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/08/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "SEP") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/09/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "OCT") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/10/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "NOV") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/11/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
  else if (substr(covData1$Appointment_Date[i], 3, 5) == "DEC") {
    covData1$Appointment_Date[i] <- paste0(substr(covData1$Appointment_Date[i], 0, 2), "/12/", substr(covData1$Appointment_Date[i], 6, 9))
    
  }
}

covData1$Weekday <- wday(dmy(covData1$Appointment_Date))
covData1$Weekday <- recode(covData1$Weekday, 
                             "1"="Sunday",
                             "2"="Monday",
                             "3"="Tuesday",
                             "4"="Wednesday",
                             "5"="Thursday",
                             "6"="Friday",
                             "7"="Saturday")

# summing the data over days
sumData <- aggregate(covData1$COUNT_OF_APPOINTMENTS, by=list(Appointment_Date=covData1$Appointment_Date), FUN=sum)

# unifying th date format
sumData$Weekday <- wday(dmy(sumData$Appointment_Date))
sumData$Weekday <- recode(sumData$Weekday, 
                            "1"="Sunday",
                            "2"="Monday",
                            "3"="Tuesday",
                            "4"="Wednesday",
                            "5"="Thursday",
                            "6"="Friday",
                            "7"="Saturday")

# plotting the relationship between appointments and weekday
sumData$Appointment_Date <- dmy(sumData$Appointment_Date)
ggplot(sumData, aes(x = Appointment_Date, y = x, color = Weekday)) + geom_jitter() + labs(y = "Visits per day", x = "Date")
Weekmean <- aggregate(sumData$x, by=list(a = sumData$Weekday), FUN = mean)



# importing the coverage dataset

coverage <- import("Appointments_GP_Daily_Aug19/coverage.csv")
coverage <- coverage[which(coverage$COMMISSIONER_ORGANISATION_CODE == "05A"),]

for (i in 1:length(coverage$Appointment_Month)) {
  if (substr(coverage$Appointment_Month[i], 3, 5) == "JAN"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/01/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "FEB"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/02/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "MAR"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/03/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "APR"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/04/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "MAY"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/05/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "JUN"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage17$Appointment_Month[i], 0, 2), "/06/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "JUL"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/07/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "AUG"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/08/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "SEP"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/09/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "OCT"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/10/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "NOV"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/11/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
  else if (substr(coverage$Appointment_Month[i], 3, 5) == "DEC"){
    coverage$Appointment_Month[i] <- paste0(substr(coverage$Appointment_Month[i], 0, 2), "/12/",20,  substr(coverage$Appointment_Month[i], 8, 11))
  }
}

# merging coverage with covData
covData1$patients <- list(0)
covData1$all_patients <- list(0)
covData1$ratio_patients <- list(0)


for (i in 1:nrow(coverage)) {
  for (j in 1:nrow(covData1)) {
    if(substr(coverage$Appointment_Month[i], 4, 11) == substr(covData1$Appointment_Date[j], 4, 11)) {
      covData1$patients[j] <- coverage$`Patients registered at included practices`[i]
      covData1$all_patients[j] <- coverage$`Patients registered at open practices`[i]
      
    }
    
  }
}

covData1$ratio_patients <- as.numeric(covData1$patients)/as.numeric(covData1$all_patients)


#write CovData1 in a csv file
covData1$patients <- unlist(covData1$patients)
covData1$all_patients <- unlist(covData1$all_patients)


write.csv(covData1, "covData1.csv", row.names = FALSE)
```
The files containing information about coverage  need to be merged before being binded with main dataset.
```{r coverage}
coverage_alpha <- vector(mode = "numeric", 20)
coverage_alpha <- import("Appointments_GP_Daily_Aug19/APPOINTMENTS_GP_COVERAGE.csv")
coverage_alpha <- coverage_alpha[which(coverage_alpha$COMMISSIONER_ORGANISATION_CODE == "05A"),]

coverage_a <- import("Appointments_GP_Daily_Aug19/cov_a.csv")
coverage_a <- coverage_a[which(coverage_a$COMMISSIONER_ORGANISATION_CODE == "05A"),]
coverage_alpha <- rbind(coverage_alpha, coverage_a[15:16, ])
write.csv(coverage_alpha, "coverage.csv")

coverage19 <- import("Appointments_GP_Daily_Aug19/APPOINTMENTS_GP_COVERAGE2.csv")
coverage19 <- coverage19[which(coverage19$COMMISSIONER_ORGANISATION_CODE == "05A"),]
coverage19 <- coverage19[17:18,]
coverage19 <-select(coverage19, -STPCD, -REGIONAL_LOCAL_OFFICE_CODE, -REGION_CODE)
coverage19$Appointment_Month[1] <- "01/11/2019"
coverage19$Appointment_Month[2] <- "01/12/2019"

coverage17a <- coverage17[1:4, ]
coverage17a <- select(coverage17a, -STPCD, -SUBREGION_CODE, -REGION_CODE)
names(coverage17a)[4] <- "Open Practices"
coverage18a <- select(coverage, -V1, -STPCD, -REGIONAL_LOCAL_OFFICE_CODE, -REGION_CODE)
full_coverage <- rbind(coverage17a, coverage18a, coverage19)
full_coverage[20, 2] <- "01/06/2019"
full_coverage$Appointment_Month <- dmy(full_coverage$Appointment_Month)
full_coverage$MonthYear <- substr(full_coverage$Appointment_Month, 1,7) 
full_coverage <- full_coverage[3:26, ]
```


Final merger of the datasets. 

```{r merger, message = FALSE}
data17 <- import("covData17a.csv")
data18 <- import("covData1.csv")
data17$Appointment_Date <- dmy(data17$Appointment_Date)
data18$Appointment_Date <- dmy(data18$Appointment_Date)
data17 <- filter(data17, Appointment_Date <= as.Date("2018-02-28"))

merged_coventry <- rbind(data17, data18)

merged_sumData <- aggregate(merged_coventry$COUNT_OF_APPOINTMENTS, by=list(Appointment_Date=merged_coventry$Appointment_Date), FUN=sum)
merged_sumData <-  mutate(merged_sumData, Appointment_Date = as.Date(Appointment_Date))
merged_sumData <-  complete(merged_sumData,Appointment_Date = seq.Date(min(Appointment_Date), max(Appointment_Date), by= 'day'))
merged_sumData$Weekday <- wday(merged_sumData$Appointment_Date)
merged_sumData$Weekday <- recode(merged_sumData$Weekday, 
                                 "1"="Sunday",
                                 "2"="Monday",
                                 "3"="Tuesday",
                                 "4"="Wednesday",
                                 "5"="Thursday",
                                 "6"="Friday",
                                 "7"="Saturday")

merged_sumData <- merged_sumData[62:791, ]
names(merged_sumData)[2] <- "Appointment_Count"

sum <- 0
j <-0
for (i in 1:nrow(merged_sumData)) {
  if (merged_sumData$Weekday[i] == "Sunday") {
    sum <- sum + merged_sumData$Appointment_Count[i]
    j <-j+1
  }
}

merged_sumData <-  mutate(merged_sumData, Appointment_Date = as.Date(Appointment_Date))
merged_sumData <-  complete(merged_sumData,Appointment_Date = seq.Date(min(Appointment_Date), max(Appointment_Date), by= 'day'))
mean <- round(sum/j)

for (i in 1:nrow(merged_sumData)) {
  if (is.na(merged_sumData$Appointment_Count[i])) {
    merged_sumData$Appointment_Count[i] <- 13
    }
}
```

Initial plot with the merged  data.
```{r initial plots}
#newer data from Mar18 till Aug19

#p_new <- ggplot(sumData, aes(x = Appointment_Date, y = Appointment_Count, color = Weekday)) + geom_jitter() + labs(y = "Visits per day", x = "Date") + ylim(0, 12000)

#older data from Nov17 till Oct 18

#p_old <- ggplot(sumData17, aes(x = Appointment_Date, y = Appointment_Count, color = Weekday)) + geom_jitter() + labs(y = "Visits per day", x = "Date") + ylim(0, 12000)

#all data from Nov17 till Aug19
p_all <- ggplot(merged_sumData, aes(x = Appointment_Date, y = Appointment_Count, color = Weekday)) + geom_jitter() + labs(y = "Visits per day", x = "Date") + ylim(0, 12000)
p_all
#  comparison of three distributions
#grid.arrange(p_old, p_new, p_all, nrow = 1)
```

Features such as season, month, coverage ratio are added now.
```{r adding features}
# add month feature
merged_sumData$MonthYear <- substr(merged_sumData$Appointment_Date, 1,7) 
merged_sumData$Month <- substr(merged_sumData$Appointment_Date, 6,7) 
# delete time feature
merged_coventry <-select(merged_coventry, -TIME_BETWEEN_BOOK_AND_APPT)

# add no of  patients to merged_Sumdata
merged_sumData$patients <- vector(mode = "numeric", nrow(merged_sumData))
merged_sumData$all_patients <- vector(mode = "numeric", nrow(merged_sumData))
merged_sumData$ratio_patients <- vector(mode = "numeric", nrow(merged_sumData))

#time series  with the fitted trend line before adjusting
ggplot(merged_sumData, aes(x = Appointment_Date, y = Appointment_Count)) + geom_jitter(aes(color = Weekday)) + labs(y = "Visits per day", x = "Date") + stat_smooth(method = lm)

write.csv(merged_sumData, "merged_sumData.csv")


# new way to merge coverage with mergedsumdata



for( i in 1:nrow(full_coverage)) {
  for(j in 1:nrow(merged_sumData)) {
    if(merged_sumData$patients[j]==0) {
      if (merged_sumData$MonthYear[j]==full_coverage$MonthYear[i]) {
        merged_sumData$patients[j] <- full_coverage$`Patients registered at included practices`[i]
        merged_sumData$all_patients[j] <- full_coverage$`Patients registered at open practices`[i]
      }
    }
  }
}

#adjusting the no of appointments based on ratio of included patients

merged_sumData$ratio_patients <- merged_sumData$patients/merged_sumData$all_patients
merged_sumData$adjusted_count <- round(merged_sumData$Appointment_Count/merged_sumData$ratio_patients)

# save merged_sumdata to shorten compilation time
write.csv(merged_sumData, "merged_sumData1.csv")
merged_sumData1 <-  import("merged_sumData1.csv")
merged_sumData1 <- select(merged_sumData1, -V1)
```

Exploratory Data Analysis for the Question 1
```{r eda, message = FALSE}

weekday_mean <- group_by(merged_sumData, Weekday) %>% summarise(count = mean(Appointment_Count))
weekday_mean$Weekday <- factor(weekday_mean$Weekday,levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
weekday_mean$Workday <- vector(mode = "character", nrow(weekday_mean))
for (i in 1:nrow(weekday_mean)) {
  if (weekday_mean$Weekday[i] == "Saturday" | weekday_mean$Weekday[i] == "Sunday") {
    weekday_mean$Workday[i] <- "Weekend"
  }
  else {
    weekday_mean$Workday[i] <- "Workday"
    }
}
weekday_mean$Workday <- factor(weekday_mean$Workday,levels = c("Workday", "Weekend"))


              # seasonality analysis
# add season indicator
merged_sumData$Season <- vector(mode = "character", nrow(merged_sumData))

for (i in 1:nrow(merged_sumData)) {
  if (merged_sumData$Month[i] == "12" || merged_sumData$Month[i] == "01" || merged_sumData$Month[i] == "02")  {
    merged_sumData$Season[i] <- "Winter"
  }
  if (merged_sumData$Month[i] == "03" || merged_sumData$Month[i] == "04" || merged_sumData$Month[i] == "05" )  {
    merged_sumData$Season[i] <- "Spring"
  }
  if (merged_sumData$Month[i] == "06" || merged_sumData$Month[i] == "07" || merged_sumData$Month[i] == "08")  {
    merged_sumData$Season[i] <- "Summer"
  }
  if (merged_sumData$Month[i] == "09" || merged_sumData$Month[i] == "10" || merged_sumData$Month[i] == "11")  {
    merged_sumData$Season[i] <- "Autumn"
  }
}

merged_sumData_work <- filter(merged_sumData, Weekday == "Monday" | Weekday == "Tuesday" | Weekday == "Wednesday" | Weekday == "Thursday" | Weekday ==  "Friday")
merged_sumData_work <- filter(merged_sumData_work, Appointment_Count >100)
month_mean <- group_by(merged_sumData_work, Month, Season) %>% summarise(count = mean(Appointment_Count))
month_mean$Month <- factor(month_mean$Month,levels = c("01", "02", "03", "04", "05", "06","07", "08", "09", "10", "11", "12"))
month_mean$Season <- factor(month_mean$Season,levels = c("Spring", "Summer", "Autumn", "Winter"))



# add year feature
merged_sumData$year <- year(merged_sumData$Appointment_Date)
merged_sumData$Season <- factor(merged_sumData$Season, levels = c("Spring", "Summer", "Autumn", "Winter"))
p_season <- ggplot(merged_sumData_work, aes(x = Appointment_Date, y = adjusted_count)) + geom_jitter(aes(color = Season)) + labs(y = "Visits per day", x = "Date") + ylim(0, 12000) + stat_smooth(method = "loess", se = FALSE) + scale_color_brewer(palette="Dark2")
p_season
merged_sumData_work <- merged_sumData_work[]
one_year <- merged_sumData_work[43:295, ]
p_season2 <- ggplot(one_year, aes(x = Appointment_Date, y = adjusted_count, color = Season)) + geom_point() + labs(y = "Visits per day", x = "Date") + ylim(0, 12000) + stat_smooth(method = "lm", formula = y ~ x, se = FALSE) + scale_color_brewer(palette="Dark2")
p_season2
by_season <- group_by(merged_sumData, Season, year) %>% summarise(count = sum(Appointment_Count))
by_month <- group_by(merged_sumData, Month, year) %>% summarise(count = sum(Appointment_Count))
by_season$year <- as.character(by_season$year)
by_month$year <- as.character(by_month$year)
by_month$MonthYear <- dmy(paste0("01/", by_month$Month, "/", by_month$year))

ggplot(data = by_season, aes(x = Season, y = count, fill = year)) + geom_bar(stat = "identity", position = position_dodge())+ scale_color_brewer(palette="Dark2")
ggplot(data = by_month, aes(x = MonthYear, y = count, fill = Month)) + geom_bar(stat = "identity", position = position_dodge()) + scale_fill_brewer(palette="Paired")


# plots for task 1
merged_sumData$Weekday <- factor(merged_sumData$Weekday,levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
merged_sumData_work$Weekday <- factor(merged_sumData_work$Weekday,levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

p_adjusted_work <- ggplot(merged_sumData_work, aes(x = Appointment_Date, y = adjusted_count)) + geom_jitter(aes(color = Weekday)) + labs(y = "Visits per day", x = "Date") + ylim(0, 12000) + stat_smooth(method = lm)+ scale_color_brewer(palette="Dark2")
p_adjusted_work

p_adjusted <- ggplot(merged_sumData, aes(x = Appointment_Date, y = adjusted_count)) + geom_jitter(aes(color = Season)) + labs(y = "Visits per day", x = "Date") + ylim(0, 12000)+ scale_color_brewer(palette="Dark2")
p_adjusted

p_not_adjusted <- ggplot(merged_sumData, aes(x = Appointment_Date, y = Appointment_Count)) + geom_jitter(aes(color = Weekday)) + labs(y = "Visits per day", x = "Date") + stat_smooth(method = lm)+ scale_color_brewer(palette="Dark2")
p_not_adjusted

weekday_mean_plot <- ggplot(weekday_mean, aes(x = Weekday, y = count, fill = Workday)) + geom_bar(stat = "identity", position = position_dodge())+ labs(x = "Weekday", y = "Average number of appointments")+ scale_fill_brewer(palette="Dark2")
weekday_mean_plot

month_mean_plot <- ggplot(month_mean, aes(x = Month, y = count, fill = Season)) + geom_bar(stat = "identity", position = position_dodge())+ labs(x = "Month", y = "Average number of appointments") + scale_fill_brewer(palette="Dark2")
month_mean_plot

```

Seasonal Decompsition of the Time Series
```{r stl, message = FALSE}
# seasonal trend decompposition
# weekly
rm(sum)
sum_weekly <- as.xts(merged_sumData$Appointment_Count, order.by = merged_sumData$Appointment_Date)
sum_weekly <- apply.weekly(sum_weekly, sum)
sum_weekly <- as.data.frame(sum_weekly)
names(sum_weekly)[1] <- "Appointment Count"
sum_weekly2 <- sum_weekly$`Appointment Count`
sum_weekly2 <- as.numeric(sum_weekly2)
sum_weekly.ts <- ts(sum_weekly2, start=c(2018, 1), frequency = 52)
stl_weekly <- stl(sum_weekly.ts, s.window = "periodic")
#plot.ts(sum_weekly.ts)
plot(stl_weekly)
decompose_weekly_add <- decompose(sum_weekly.ts,type = "additive")
decompose_weekly_mult <- decompose(sum_weekly.ts,type = "multiplicative")
plot(decompose_weekly_add)
plot(decompose_weekly_mult)

#montly
rm(sum)
sum_monthly <- as.xts(merged_sumData$Appointment_Count, order.by = merged_sumData$Appointment_Date)
sum_monthly <- apply.monthly(sum_monthly, FUN = sum)
sum_monthly <- as.data.frame(sum_monthly)

names(sum_monthly)[1] <- "Appointment Count"
sum_monthly2 <- sum_monthly$`Appointment Count`
sum_monthly2 <- as.numeric(sum_monthly2)
sum_monthly2[25] <- 187706
sum_monthly.ts <- ts(sum_monthly2, start=c(2018, 1), frequency = 12)
plot.ts(sum_monthly.ts)
stl_monthly <- stl(sum_monthly.ts, s.window = "periodic")
plot(stl_monthly)
decompose_monthly_add <- decompose(sum_monthly.ts,type = "additive")
decompose_monthly_mult <- decompose(sum_monthly.ts,type = "multiplicative")
plot(decompose_monthly_mult)


```
Task 2


Importing and merging weather data. 
```{r task2 merger}
temp <- import("temp.xlsx")
rainfall <- import("rainfall.xlsx")
temp <- temp[37:60,2]
rainfall <- rainfall[37:60,2]
monthly_sumdata <- aggregate(merged_sumData$adjusted_count, by=list(monthYear = merged_sumData$MonthYear), FUN = sum)
monthly_sumdata <- cbind(monthly_sumdata, rainfall, temp)
names(monthly_sumdata)[2] <- "Appointment_Count"
monthly_sumdata$monthYear <-paste0(monthly_sumdata$monthYear, "-01")
monthly_sumdata$monthYear <- ymd(monthly_sumdata$monthYear)
monthly_sumdata$Month <- month(monthly_sumdata$monthYear, label = TRUE)


monthly_sumdata$Season <- vector(mode = "character", nrow(monthly_sumdata))
monthly_sumdata$Season_numeric <- vector(mode = "numeric", nrow(monthly_sumdata))

for (i in 1:nrow(monthly_sumdata)) {
  if (monthly_sumdata$Month[i] == "Dec" || monthly_sumdata$Month[i] == "Jan" || monthly_sumdata$Month[i] == "Feb")  {
    monthly_sumdata$Season[i] <- "Winter"
    monthly_sumdata$Season_numeric[i] <- 4
    
  }
  if (monthly_sumdata$Month[i] == "Mar" || monthly_sumdata$Month[i] == "Apr" || monthly_sumdata$Month[i] == "May")  {
    monthly_sumdata$Season[i] <- "Spring"
    monthly_sumdata$Season_numeric[i] <- 1
    
  }
  if (monthly_sumdata$Month[i] == "Jun" || monthly_sumdata$Month[i] == "Jul" || monthly_sumdata$Month[i] == "Aug")  {
    monthly_sumdata$Season[i] <- "Summer"
    monthly_sumdata$Season_numeric[i] <- 2
    
  }
  if (monthly_sumdata$Month[i] == "Sep" || monthly_sumdata$Month[i] == "Oct" || monthly_sumdata$Month[i] == "Nov")  {
    monthly_sumdata$Season[i] <- "Autumn"
    monthly_sumdata$Season_numeric[i] <- 3
    
  }
}
```
Exploratory Data Analysis for Question 2
```{r eda2}
ggplot(monthly_sumdata, aes(x = monthYear, y = Appointment_Count)) + geom_point(aes(size = rainfall, color = temp)) + scale_color_gradient(low = "blue", high = "red")


names(monthly_sumdata)[3] <- "Rainfall"
names(monthly_sumdata)[4] <- "Temperature"
ggplot(monthly_sumdata, aes(x = temp, y = Appointment_Count)) + stat_smooth(method = "lm") + labs(x = "Average monthly temperature in Coventry in the UK(in Celsius)", y = "Total number of appointments")
ggplot(monthly_sumdata, aes(x = rainfall, y = Appointment_Count)) + stat_smooth(method = "lm")+ labs(x = "Total monthly rainfall in the UK(in mm)", y = "Total number of appointments")


monthly_sumdata$Month <- month(monthly_sumdata$monthYear, label = FALSE)
corrdata <- select(monthly_sumdata, Appointment_Count, Rainfall, Temperature)
names(corrdata)[1] <- "Number of appointments"


corelation <- cor(corrdata)
ggcorrplot(corelation, hc.order = FALSE, type = "full", lab = TRUE) 
```